\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2016-07-25 redaction/dhp_draft_july2016.tex
Modifications Guillaume

\documentclass[a4paper,11pt]{article}
%\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
%\usepackage[latin1]{inputenc}
%\usepackage{hyperref}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{lscape}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{vmargin}
\usepackage{epstopdf}
\usepackage{hyperref}
\usepackage{float}
\usepackage{caption}
\usepackage[tableposition=top]{caption}
\usepackage{ bbold }
\usepackage{ upgreek }
%\usepackage[nolists, figuresfirst]{endfloat}
%\usepackage{sidefloat}

@@

\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0,5cm}{0,5cm}{0,3cm}{0,5cm}

@@

\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}} %\thanks{We are especially grateful to XXX. Part of this research was funded by the French Agence Nationale de la Recherche (ANR), under grant ANR-11-JSH1 002 01. Finally, any remaining errors are ours.}}}
%We also thank participants at the EEA 27\textsuperscript{th} Congress (Oslo), ETSG 13\textsuperscript{th} meetings (Copenhagen), INFER Workshop (Orleans), 45\textsuperscript{th} Congress of the Canadian Economic Association (Ottawa), IAW Workshop (Tubingen), and seminars at WTO, Strathclyde University, OFCE and EQUIPPE-University of Lille.
\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris Dauphine, LEDa \& PSL, Chercheur associ\'{e} - Sciences Po (OFCE)}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS (UMR 9221), Universit\'{e} de Bretagne Occidentale-ICI (EA 2652) and CEPII}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris Dauphine, LEDa \& PSL} }
 \maketitle

@@

\begin{abstract}
Following \cite{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an ad valorem tax equivalent. However, many common empirical facts support the existence of additive costs. This paper contributes to this debate. Precisely, we provide a quantitative assessment of the size and the importance of the additive component in international transport costs. Using SITC 3 and 4 digit cif-fob unit values over 1974-2013 taken from US imports data, we estimate the two components of transport costs, by transport mode (air or ocean). We find that additive costs are 2.85\% of fob unit values for ocean shipping (and ad-valorem ones 3.22\%). These values are respectively equal to 1.9\% and 2.5\% for air transport. Further, we show that taking additive costs into account improves the fit of the modelling of transport costs. We also use the time dimension of our data to characterize the evolution of transport costs. After correcting for composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013. Yet, this steady decline hides shifts in the relative importance of additive and ad-valorem. While most of the early decline in air transport costs can be explained by the ad-valorem component, this component nearly doubled in the 2000s. These results therefore confirm the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

\noindent Trade costs remain central in international economic analysis. In particular, they are considered as a major obstacle to international economic integration and international trade flows. According to the estimates by \citet{Jacks08}, trade cost declines explain around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse. After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO starting in 1995) multilateral agreements. Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000. Yet, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs (typically captured through distance) still remain a major obstacle to trade (e.g. \citealp{Head_Mayer04} and \citealp{Disdier_Head08}). Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries put up to 74\% markups over production costs. Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}. In this vein, \citet{anderson_wincoop_jel} obtain that the 74\% markup made by international trade costs can be divided in 44\% for border-related trade barriers and 21\% for transportation costs. This last figure suggests that the share of transport costs in the consumer price of manufactured goods is substantial. \citet{Behar_Venables} obtain that the elasticity of trade with respect to freight costs is sizeable, by around -3. If much of trade-policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention. This is accordingly the focus of the paper.\smallskip

Following \citet{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an \emph{ad valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, the ``iceberg cost'' hypothesis). However, many common empirical facts support the existence of additive costs. As documented by \citet{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs... often exhibit (at least partly) an additive structure. The structure (additive vs multiplicative) of transport costs is far from being anecdotal, as the literature has long emphasized its role in shaping the pattern of trade flows. The Alchian and Allen conjecture (\citealp{alchian}), which points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: The relative demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out''). \citet{Lashkaripour-mimeo-2016} challenges this view though. He finds supporting evidence for the iceberg assumption by taking into account the fact that more expensive goods are systematically heavier and hence more costly to transport. One can yet be concerned by the generality of his results though. By nature, his study is restricted to goods that are enumerated by items in the statistics (that represent 60\% of US imports for instance). Furthermore, while the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study). Besides, a number of empirical papers provide a strong empirical support to the role of additive costs in international. Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher fob unit values on exports to more remote countries, in contradiction with the iceberg hypothesis. \citet{hummels_skiba} estimate the elasticity of freight rates with respect to price is estimated to be well below unity. Also, their estimates imply that doubling freight costs increases average fob export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs. These findings deliver strong empirical support in favor of the Alchian-Allen conjecture. The existence of additive costs may also explain a large number of zeros in bilateral trade flows, and more generally, the granularity of trade flows. Relying on Spanish and US transaction-level trade data, \citet{Hornok14} find that additive trade costs of are associated with less frequent and larger shipments. With additive costs, exporters wait to fill completely a container before sending it abroad, to decrease as much as possible the number of shipments, thereby inducing more ``lumpiness'' in international trade. \smallskip

Beyond the positive aspect of understanding trade patterns, several recent papers also point out the normative implications of additive trade costs. \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the iceberg component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in multiplicative costs. Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized multiplicative tariff. While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much can be done in quantifying such gains though, one potential reason being the lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to palliate this gap. \bigskip

Our paper contributes to the literature by providing results on the size of transport costs over time, explicitly distinguishing between multiplicative and additive parts. Precisely, we quantitatively assess the size and the importance of the additive component in international transport costs. To do so, we update the detailed US customs, sector-level data from the US Imports of Merchandise used by \citet{hummels2007}, to cover the period from 1974 to 2013. Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian exports in 2004, their results suggest that additive costs are about 14\% of the median consumer price. However, while our data requires that we only consider transportation costs, our approach departs from theirs in several key aspects, as we develop below. One important difference is that we exploit exhaustive information about the imports flows of the US, over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. They can be summarized in three main points.

First, our theoretically agnostic approach provides a fairly simple framework for measuring both multiplicative and additive parts of transportation costs, which represents a valuable insight for calibrating related models.\footnote{Given their database and their methodology, \citet{Irrazabal_2015} can only identify the ratio of the additive to the multiplicative cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median fob price by country-product ($\bar{\widetilde{p}}_{ik}$). By contrast, our estimation strategy enables us to uncover both values of the iceberg and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product fob price $\widetilde{p}_{ik}$.} We thus obtain that the mean values over 1974-2013 of iceberg costs are equal to 2.5\% and 3.2\% in air and ocean respectively, whereas the additive component amounts to 1.8\% and 2.9\% of the fob price (in air and ocean). To our best knowledge, our paper is the first to provide such an extensive quantitative measure of both multiplicative and additive costs in total transport costs. Second, we provide an empirical assessment of what standard international trade models lose by skipping additive transport costs. Quantitatively, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2. On average over the whole period, biased estimate for iceberg is 5\% for Air and 6\% for Vessel, while unbiased estimate is respectively 2.5\% and 3.2\%. We also show that taking additive costs into account significantly improves the fit of the modelling of transport costs, through various measures of ``goodness-of-fit'', for both transport modes and years.

Third, we exploit the time dimension of our database to document the patterns of transport costs over time. Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see\citealp{Lafourcade_Thisse}). \cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms. However, \cite{hummels2007} shows that the bulk of price declines in transportation comes from air transport, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004. Concerning ocean shipping, which represents the major part of world trade value, decline in trade prices are much less obvious, a conclusion in accordance with the studies reviewed by \cite{Behar_Venables}. Our paper contributes to this debate. As pointed out by \cite{hummels2007}, we confirm the importance of the composition effects by country of origin and by product in shaping the patterns of transportation costs over time. After excluding the composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013. Yet, this steady decline hides shifts in the relative importance of additive and ad-valorem. Air transport in particular exhibit contrasted trends. While most of the early decline in air transport costs can be explained by the ad-valorem part, this component nearly doubled in the 2000s, all along with a substantial decline in the per-unit cost. These results therefore confirm the importance of the additive component in accounting for international transport costs.\smallskip

The paper is built as follows. \ref{sec:data_method} explains the data sources and the empirical methodology retained in the paper. Sections \ref{sec:results_decomposition} and \ref{sec:results_trends} report our results. In Section \ref{sec:results_decomposition}, we characterize the role of the additive component of transport costs. After reporting the mean values over the period (by transport mode), we show the improved performance in including additive trade costs in the measure of transport costs. This being established, Section \ref{sec:results_trends} characterizes the trends in each component of transport costs (by transport mode) over the period 1974-2012. Section \ref{sec:conclu} concludes.

@@

As in \cite{hummels2007} (among others), our measure of transportation costs consists in exploiting the difference between commodity-level export and import prices.
We first use values, quantities and freight costs to recover free-on-board (FOB) and cost-insurance-fret (CIF) prices, by goods, country of origin and transportation mode. More precisely, the (unit) FOB price is computed as the total customs value divided by the shipping weight; in other words, it is the price for the good net of transportation costs. The CIF price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight. Our dependant variable is finally computed as the ratio of the CIF price divided by the FOB price. Higher than 1, the variable provides therefore with a measure of transport costs as a proportion of the good's price, an \emph{ad valorem} equivalent. The database we use to construct our measure of transport comes from US annual Imports of Merchandise provided by the Census bureau\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}}, spanning from 1974 to 2013. It is worth acknowledging that using this dataset encounters some limitations. First, our measure of the cif-fob price gap only covers transportation costs by nature, thereby being silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.  Further, in terms of transport costs \textit{per se}, this measure being based on freight costs, omits the other dimension of transport costs related to the time value of goods in transit. According to \citet{anderson_wincoop_jel}), the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit. In this respect, it is true that using this dataset embraces a partial view of international transport barriers.

Yet, we believe that the results we get deliver valuable insights on this topic. Indeed, using this dataset has (at least) three main advantages. First, this dataset delivers a strong statistical reliability arising from a single, trustworthy customs origin. Based on customs declarations, this dataset inventories all imports (both values and quantities) by country of origin to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system. In addition, the database reports information regarding freight expenditures and transportation mode (Ocean vessel and Air). The first will be crucial to compute transport costs (see below), the second will allow us enlightening substantial differences in the dynamics of transport costs across transportation mode. Second, using this dataset allows us to have the import price of the good (CIF price), next to the export price (FOB price). This is highly valuable, as we can estimate both the \textit{levels} of the iceberg trade costs and of the additive trade costs. This differentiates us from \citet{Irrazabal_2015}, which can only estimate the ratio of additive costs as a share of the median total consumer price. Third, this dataset having been used by \cite{hummels2007}, this enables us to compare our results to his findings. Further, we extend the time coverage to the more recent period up to 2013, while \cite{hummels2007} stops in 2004. As we show in Section \ref{sec:results_trends}, covering the time period over the recent years delivers interesting insights regarding the trends in air shipping costs.

We make use sectorial price data at the 3-digit level, even if the data of transportation costs is available at a more disaggregated level. This is primarily due to technical reasons. As detailed below, the use of a nonlinear estimator triggers computational limitations that do not make them a likely option, especially when covering a long period of time. Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level (for some selected years).\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.} Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation. Depending on the considered year, this leaves us with around 200 (3-digits) and 600-700 (4-digits) products, from approximately 200 countries of origin.

@@

\paragraph{The estimated equation} Our purpose is to provide estimates over time of the sizes of multiplicative and additive costs among total transport costs. To do so, we start from the equation that expresses the price paid by the consumer (import, or cif) price $p$ as a function of the producer (export, or fob) price $\widetilde{p}$ given both per-unit ($t$) and ad-valorem ($\tau$) transport costs, according to:\footnote{This equation is similar to \cite{Irrazabal_2015} or \cite{martin2012}, among others.}

@@

\noindent As usual in the literature, the so-called ``iceberg'' trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no iceberg trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive trade costs). We estimate this equation for each year over the period 1974-2013, and for each of the two transportation modes reported (air or vessel), on a sectoral-origin country basis.\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.} Let us denote $i$ the origin country, and $k$, the sector (or product). Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following equation at the root of our estimation:

@@

\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1°)} both multiplicative and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2)} in a multiplicative way for the former and an additive way for the latter. In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:

\begin{eqnarray}
\tau_{ik} &=& \tau_{i} \times \tau_{k} \label{eq:iceberg}\\
t_{ik} &=& t_{i} + t_{k} \label{eq:add}
\end{eqnarray}

@@

\subsection{Fob-cif prices}

>>>>

\documentclass[a4paper,11pt]{article}
%\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
%\usepackage[latin1]{inputenc}
%\usepackage{hyperref}
%\usepackage[T1]{fontenc}
%\usepackage[utf8]{inputenc}
%%GD Quand je mets usepackage[utf8]{inputenc} je n'arrive plus à compiler la biblio
%J'ai essayé de débuger : je n'y suis pas arrivé
\usepackage{rotating}
\usepackage{setspace}
\usepackage{lscape}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{vmargin}
\usepackage{epstopdf}
\usepackage{hyperref}
\usepackage{float}
\usepackage{caption}
\usepackage[tableposition=top]{caption}
\usepackage{ bbold }
\usepackage{ upgreek }
\usepackage{comment}
\includecomment{commentGD}
\usepackage[draft]{todonotes}
%%GDPour cacher les notes, \usepackage[disable]{todonotes}

%\usepackage[nolists, figuresfirst]{endfloat}
%\usepackage{sidefloat}

@@

\setmarginsrb{3cm}{2cm}{3cm}{2cm}{0,5cm}{0,5cm}{0,3cm}{0,5cm}

@@

\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}} %\thanks{We are especially grateful to XXX. Part of this research was funded by the French Agence Nationale de la Recherche (ANR), under grant ANR-11-JSH1 002 01. Finally, any remaining errors are ours.}}}
%We also thank participants at the EEA 27\textsuperscript{th} Congress (Oslo), ETSG 13\textsuperscript{th} meetings (Copenhagen), INFER Workshop (Orleans), 45\textsuperscript{th} Congress of the Canadian Economic Association (Ottawa), IAW Workshop (Tubingen), and seminars at WTO, Strathclyde University, OFCE and EQUIPPE-University of Lille.
\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE \newline
SciencesPo, Observatoire Français des Conjonctures \'{E}conomiques (OFCE), 75007, PARIS, FRANCE}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS (UMR 9221), Universit\'{e} de Bretagne Occidentale-ICI (EA 2652) and CEPII}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE \newline
 } }
 \maketitle

@@

\begin{abstract}
Following \cite{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an ad-valorem tax equivalent. However, many common empirical facts support the existence of additive costs. This paper provides a quantitative assessment of the size and the importance of the additive component in international transport costs.
Using SITC 3 and 4 digit cif-fab unit values over 1974-2013 taken from US imports data, we estimate the two components of transport costs by transport mode (air or ocean).
We find that additive costs are 2.85\% of fab unit values for ocean shipping (and ad-valorem ones 3.22\%). These values are respectively equal to 1.9\% and 2.5\% for air transport. Further, we show that taking additive costs into account improves the fit of the modelling of transport costs.
We also use the time dimension of our data to characterize the evolution of transport costs. After correcting for composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013. Yet, this steady decline hides shifts in the relative importance of additive and ad-valorem. While most of the early decline in air transport costs can be explained by the ad-valorem component, this component nearly doubled in the 2000s. These results therefore confirm the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

\noindent Trade costs are central in international economic analysis. In particular, they are considered a major obstacle to international economic integration and international trade flows. According to \citet{Jacks08}, trade cost declines explain around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse. After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO starting in 1995) multilateral agreements.
Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.
Yet, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs (typically captured through distance) still remain a major obstacle to trade (e.g. \citealp{Head_Mayer04} and \citealp{Disdier_Head08}). Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.
Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}. In this vein, \citet{anderson_wincoop_jel} obtain that the 74\% markup made by international trade costs can be divided between 44\% for border-related trade barriers and 21\% for transportation costs. This last figure suggests that the share of transport costs in the consumer price of manufactured goods is substantial.
\citet{Behar_Venables} obtain that the elasticity of trade with respect to freight costs is sizeable, by around -3.
\todo{GD 230716 Euh... je crois que c'est un non-sequitur, non ?}
If much of trade-policy barriers have been removed over the second half of the 20th century, these findings point out that the transport costs component of the overall trade costs remain large and deserve attention. This is accordingly the focus of the paper.\smallskip

Following \citet{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).
\todo{GD 230716 iceberg, c'est deux choses : ad-valorem + le fait que les coûts sont payés avec des bouts du bien transporté}
However, many common empirical facts support the existence of additive costs. As documented by \citet{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs... often exhibit (at least partly) an additive structure. The structure (additive vs ad-valorem) of transport costs is far from being anecdotal, as the literature has long emphasized its role in shaping the pattern of trade flows. The Alchian and Allen conjecture (\citealp{alchian}), points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: The relative demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out'').
\citet{Lashkaripour-mimeo-2016} challenges this view. He finds supporting evidence for the \emph{ad-valorem} assumption by taking into account the fact that more expensive goods are systematically heavier and hence more costly to transport. One can yet be concerned by the generality of his results though. By nature, his study is restricted to goods that are enumerated by items in the statistics (they represent 60\% of US imports). Furthermore, while the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour's study).
Besides, a number of empirical papers provide a strong empirical support to the role of additive costs in international. Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher fab unit values on exports to more remote countries, in contradiction with the ad-valorem hypothesis. \citet{hummels_skiba} estimate the elasticity of freight rates with respect to price is estimated to be well below unity. Also, their estimates imply that doubling freight costs increases average fab export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs. These findings deliver strong empirical support in favor of the Alchian-Allen conjecture.
The existence of additive costs may also explain a large number of zeros in bilateral trade flows, and more generally, the granularity of trade flows. Relying on Spanish and US transaction-level trade data, \citet{Hornok14} find that additive trade costs of are associated with less frequent and larger shipments. With additive costs, exporters wait to fill completely a container before sending it abroad, to decrease as much as possible the number of shipments, thereby inducing more ``lumpiness'' in international trade.
\todo{GD 230716 pas sur que cela soit le sujet. Plutôt «per shipment» / «fixed costs» que «additive»} \smallskip

Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs. \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs. Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff. While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains. One potential reason is the lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to palliate this gap. \bigskip

Our paper contributes to the literature by providing results on the size of transport costs in long run, explicitly distinguishing between ad-valorem and additive parts. Precisely, we quantitatively assess the size and the importance of the additive component in international transport costs. To do so, we update the detailed US customs, sector-level data from the US Imports of Merchandise used by \citet{hummels2007}, to cover the period from 1974 to 2013. Closely related to our paper is the work by \citet{Irrazabal_2015}, that develops a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian exports in 2004, this work finds that additive costs are about 14\% of the median consumer price. However our approach departs from theirs in several key aspects, as we develop below. Our data are restricted to transport, insurance and handling costs. Still, we exploit exhaustive information about the imports flows of the US, over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. They can be summarized in three main points.

First, our theoretically agnostic approach provides a fairly simple framework for measuring both ad-valorem and additive parts of transportation costs. This represents a valuable insight for calibrating related models.\footnote{Given their database and their methodology, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median fab price by country-product ($\bar{\widetilde{p}}_{ik}$). By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product fab price $\widetilde{p}_{ik}$.} We thus obtain that the mean values over 1974-2013 of ad-valorem costs are equal to 2.5\% and 3.2\% of the fab price in air and ocean transport respectively, whereas the additive component amounts to 1.8\% and 2.9\% . To our best knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.

Second, we provide an empirical assessment of what standard international trade models lose by skipping additive transport costs. Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2. On average over the whole period, biased estimate for ad-valorem is 5\% for air and 6\% for vessel, while unbiased estimate is respectively 2.5\% and 3.2\%. We also show that taking additive costs into account significantly improves the fit of the modelling of transport costs, through various measures of ``goodness-of-fit'', for both transport modes and years.

Third, we exploit the time dimension of our database to document the patterns of transport costs over time. Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see\citealp{Lafourcade_Thisse}). \cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms. However, \cite{hummels2007} shows that the bulk of price declines in transportation comes from air transport, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004. Concerning ocean shipping, which represents the major part of world trade, decline in trade prices are much less obvious, a conclusion in accordance with the studies reviewed by \cite{Behar_Venables}. Our paper contributes to this debate. As pointed out by \cite{hummels2007}, we confirm the importance of the composition effects by country of origin and by product in shaping the patterns of transportation costs over time. After excluding the composition effects, we find that all types of transport costs have been roughly constant from 1974 to 1984 and then steadily decreased by 40\% over the period 1984-2013. Yet, this steady decline hides shifts in the relative importance of additive and ad-valorem costs. Air transport in particular exhibit contrasted trends. While most of the early decline in air transport costs can be explained by ad-valorem costs, this component nearly doubled in the 2000s, all along with a substantial decline in the per-kg cost. These results therefore confirm the importance of the additive component in accounting for international transport costs.\smallskip

The paper is built as follows. \ref{sec:data_method} presents our data and our  methodology. Sections \ref{sec:results_decomposition} and \ref{sec:results_trends} report our results. In Section \ref{sec:results_decomposition}, we characterize the role of the additive component of transport costs. After reporting the mean values over the period (by transport mode), we show how introducing additive transport costs improves the modelling of transport costs. Section \ref{sec:results_trends} characterizes the trends in each component of transport costs (by transport mode) over the period 1974-2013. Section \ref{sec:conclu} concludes.

@@

As in \cite{hummels2007} (among others), our measure of transportation costs consists in exploiting the difference between commodity-level export and import prices.
We first use customs values, quantities and freight costs to recover free-alongside-board (FAB) and cost-insurance-fret (CIF) prices, by goods, country of origin and transportation mode. More precisely, the (unit) FAB price is computed as the total "customs value" in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs. The CIF price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight. Our dependant variable is finally computed as the ratio of the CIF price divided by the FAB price. Higher than 1, the variable provides therefore with a measure of transport costs as a proportion of the good's price, an \emph{ad-valorem} equivalent. The database we use to construct our measure of transport comes from US annual Imports of Merchandise provided by the Census bureau\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}}, spanning from 1974 to 2013. It is worth acknowledging that using this dataset encounters some limitations. First, our measure of the cif-fab price gap only covers freight, insurance and handling costs. It is silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.  Further, in terms of transport costs \textit{per se}, this measure being based on freight costs, omits the other dimension of transport costs related to the time value of goods in transit. According to \citet{anderson_wincoop_jel}), the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit. In this respect, it is true that using this dataset embraces a partial view of international transport barriers.

Yet, we believe that our deliver valuable insights on this topic. Indeed, using this dataset has (at least) three main advantages. First, this dataset delivers a strong statistical reliability arising from a single, trustworthy, customs origin. Based on customs declarations, this dataset inventories all imports (both values and quantities) by country of origin to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system. In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode (ocean (or "vessel") and air). The first will be crucial to compute transport costs (see below), the second will allow to enlight substantial differences in the dynamics of transport costs across transportation mode. Second, using this dataset allows us to have the import price of the good (CIF price), next to the export price (FAB price). This is highly valuable, as we can estimate both the \textit{levels} of the ad-valorem trade costs and of the additive trade costs. This differentiates us from \citet{Irrazabal_2015}, who can only estimate the ratio of additive costs as a share of the median total consumer price. Third, this dataset has been used by \cite{hummels2007}, this enables us to compare our results to his findings. Further, we extend the time coverage to the more recent period up to 2013, while \cite{hummels2007} stops in 2004. As we show in Section \ref{sec:results_trends}, covering the time period over the recent years delivers interesting insights regarding the trends in air shipping costs.

We use sectorial price data at the 3-digit level, even if the data of transportation costs is available at a more disaggregated level. As detailed below, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time. Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013.} Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation. Depending on the considered year, this leaves us with around 200 (3-digits) and 600-700 (4-digits) products, from approximately 200 countries of origin.

@@

\paragraph{The estimated equation} Our purpose is to provide estimates over time of the size of ad-valorem and additive costs among total transport costs. To do so, we start from the equation that expresses the price $p$ paid by the importer
\todo{GD 240716 importer plutôt que consumer : il reste des coûts entre les deux...}
 (import, or cif) a function of the producer price $\widetilde{p}$ (export, or fab), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:\footnote{This equation is similar to \cite{Irrazabal_2015} or \cite{martin2012}, among others.}

@@

\noindent As usual in the literature, the so-called ``ad-valorem'' trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive trade costs).  Let us denote $i$ the origin country, and $k$, the sector (or product). Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following equation at the root of our estimation\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.}:

@@

We estimate this equation for each year over the period 1974-2013, and for each of the two transportation modes reported (air or vessel), on a sectoral-origin country basis.

\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2)} this separability is in a multiplicative way for the former and an additive way for the latter. In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:

\begin{eqnarray}
\tau_{ik} &=& \tau_{i} \times \tau_{k} \label{eq:ad-valorem}\\
t_{ik} &=& t_{i} + t_{k} \label{eq:add}
\end{eqnarray}

@@

\subsection{Fab-cif prices}
